Ant cuticle image classification using texture analysis: a comparative study

### 2022-06-17 paper/main.tex
post meeting updates

\colorbox{green}{TODO: how do we connect this section to image texture analysis?
    (why do we research this and why image texture) }

\subsection{Texture Analysis}
\colorbox{green}{TODO: A brief survey on image texture and cite Liu's
    IJCV2019 survey paper}

\subsection{Deep Learning for Texture Analysis}

In the past decade, deep learning using convolutional neural networks (CNNs)
\cite{krizhevsky_imagenet_2017} has emerged as the state-of-the-art technology
for image analysis. Following this trend, various CNN-based network
architectures have been designed to specifically characterize texture images.
Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et al.
\cite{cimpoi_describing_2014} is widely accepted as one of the most important
pioneering works. It applies FV pooling to deep features obtained via a CNN
pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded
features for texture classification. Although it is capable of achieving much
improved classification accuracy comparing to traditional hand-crafted texture
features, FV-CNN does not support an end-to-end learning, where feature
extraction, encoding, and classification are separated from each other. To
achieve an end-to-end learning, Zhang et al. \cite{zhang_visual_2018} proposed
the deep texture encoding network (DeepTEN), in which a novel texture encoding
layer is added to a standard CNN architecture. Then, Xue et al.
\cite{xue_deep_2018} constructed the deep encoding pooling network (DEP), which
improves over DeepTEN by integrating local spatial characteristics into the
texture representation. Based on DeepTEN and DEP, Hu et al. \cite{hu_multi_2019}
further developed the multi-level texture encoding and representation (MuLTER)
network, which embeds a learnable encoding module at each convolutional layer so
that encoding is performed for both low-level and high-level features, yielding
a multi-level texture representation.

@@

\colorbox{green}{TODO: kviews, gabor, etc. will go here}

@@

@@

>>>>

\colorbox{green}{\parbox{\dimexpr\textwidth-2\fboxsep}{TODO: how do we connect
    this section to image texture analysis? (why do we research this and why
    image texture) }}

\subsection{Texture Analysis}
Image texture analysis is an important part of image interpretation as many
objects appearing in the textural patterns either partially or completely inside
an image. This phenomenon appears frequently in many images we encountered in
our environment as almost everybody is using the iPhone to capture the photos.
Scientists are also studying their subjects depending on taking the images
remotely such as remotely sensed satellite images. With an abundant number of
textural images surrounding us, it is essential to have a reliable automatic
tool for the textural image interpretation.

Classical computer methods are developed for textural images based on the
spatial relationship of gray levels of pixels encoded in the texture
\colorbox{green}{[Haralick and Shapiro, Hung]}. Those methods by de-facto follow
the steps used in the statistical pattern recognition
\colorbox{green}{[Fukunaga]}. Therefore, textural features are extracted, and
the classification and segmentation methods are used for the interpretation
either using the supervised or unsupervised approaches. The interpretation
method is usually taken and modified from the traditional pattern recognition
methods. For example, the gray-level co-occurrence matrix (GLCM) and local
binary patterns (LBP) are two of methods for extracting features
\colorbox{green}{[references]}. Some other approaches also have been developed
for the textural feature extraction \colorbox{green}{[references]}.

While most feature vectors are established based on a single pixel, Liu et al.
gave a comprehensive survey on textural characterization in which they call this
feature extraction as texture representation \colorbox{green}{[Liu]}. Their
survey on Bag of Words and Convolutional Neural Networks are the spatial
relationship concepts. The K-views was developed for taking the spatial
information in the clustering approach \colorbox{green}{[Hung]}.  In the past
decade, the deep learning network has become the mainstream in image
classification and segmentation \colorbox{green}{[references]}. Compared with
the traditional method in which a kernel must be designed by an engineer for
extracting features, the deep learning network can automatically extract the
features through the training. In addition, the deep learning networks achieve
the higher accuracy than that of the classical approaches although the deep
networks require much more data for training.

Similar to the traditional pattern recognition, classical image texture analysis
methods can be grouped into four categories: statistical, structural,
model-based, and transform-based methods \colorbox{green}{[references]}. Among
all of the categories, the statistical method is often used by employing the
method from the pattern recognition principle.  In doing so, an image texture is
represented as feature vectors and then fed into the algorithm. Markov random
field models are also studied for textural image interpretation
\colorbox{green}{[references]}. The transform-based methods use some functions
to decompose an image texture into a set of basic feature images. Gabor filters
and Wavelet expansions are two of the widely used approaches
\colorbox{green}{[references]}.

\colorbox{green}{TODO: add reference for sun phd thesis (Robust texture classification)}

\colorbox{green}{TODO: cleanup transition}

Recently, the deep learning network becomes the main stream in the
interpretation of textural images [Sun and liu]. In the past decade, deep
learning using convolutional neural networks (CNNs)
\cite{krizhevsky_imagenet_2017} has emerged as the state-of-the-art technology
for image analysis. Following this trend, various CNN-based network
architectures have been designed to specifically characterize texture images.
Among these, the Fisher-vector CNN descriptor (FV-CNN) proposed by Cimpoi et al.
\cite{cimpoi_describing_2014} is widely accepted as one of the most important
pioneering works. It applies FV pooling to deep features obtained via a CNN
pre-trained using the ImageNet \cite{krizhevsky_imagenet_2017} to obtain encoded
features for texture classification. Although it is capable of achieving much
improved classification accuracy comparing to traditional hand-crafted texture
features, FV-CNN does not support an end-to-end learning, where feature
extraction, encoding, and classification are separated from each other. To
achieve an end-to-end learning, Zhang et al. \cite{zhang_visual_2018} proposed
the deep texture encoding network (DeepTEN), in which a novel texture encoding
layer is added to a standard CNN architecture. Then, Xue et al.
\cite{xue_deep_2018} constructed the deep encoding pooling network (DEP), which
improves over DeepTEN by integrating local spatial characteristics into the
texture representation. Based on DeepTEN and DEP, Hu et al. \cite{hu_multi_2019}
further developed the multi-level texture encoding and representation (MuLTER)
network, which embeds a learnable encoding module at each convolutional layer so
that encoding is performed for both low-level and high-level features, yielding
a multi-level texture representation.

@@

\colorbox{green}{TODO: kviews, kmeans, etc. will go here}

@@

\colorbox{green}{TODO: Fix tables, decide on how to group models}

\begin{table}[h]
    \centering
    \caption{Results}
    \begin{tabular}{lllllll}
        \toprule
        {} & model                     & accuracy & precision & recall & f1   & roc\_auc \\
        \midrule
        0  & drp\_multi\_not\_pt       & 0.87     & 0.89      & 0.92   & 0.91 & 0.86     \\
        1  & drp\_multi\_pt            & 0.88     & 0.89      & 0.93   & 0.91 & 0.87     \\
        2  & drp\_single\_not\_pt      & 0.88     & 0.9       & 0.94   & 0.92 & 0.88     \\
        3  & drp\_single\_pt           & 0.88     & 0.9       & 0.93   & 0.91 & 0.87     \\
        4  & drp\_single\_aux\_not\_pt & 0.87     & 0.89      & 0.92   & 0.9  & 0.86     \\
        5  & drp\_single\_aux\_pt      & 0.89     & 0.9       & 0.94   & 0.92 & 0.89     \\
        6  & kviews\_17                & 0.62     & 0.79      & 0.59   & 0.68 & 0.62     \\
        7  & kviews\_19                & 0.62     & 0.8       & 0.59   & 0.68 & 0.62     \\
        8  & kviews\_25                & 0.52     & 0.71      & 0.51   & 0.59 & 0.52     \\
        9  & kmeans                    & 0.47     & 0.07      & 0.35   & 0.11 & 0.47     \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{Clustering Results}
    \begin{tabular}{llll}
        \toprule
        {} & model      & rand\_score & adjusted\_rand\_score \\
        \midrule
        0  & kviews\_17 & 0.53        & 0.06                  \\
        1  & kviews\_19 & 0.53        & 0.06                  \\
        2  & kviews\_25 & 0.5         & -0.0                  \\
        3  & kmeans     & 0.5         & 0.0                   \\
        \bottomrule
    \end{tabular}
\end{table}

@@

\colorbox{green}{TODO: mention how future work may include extra classes}
